GPU-Beschleunigung der Material Point Method

### 2019-01-25 mscthesis.tex
optimization chapter finished

To our knowledge there is no way to find out the exact number of registers an OpenGL shader uses. The program object assembly output of \textit{nvemulate} \cite{NVIDIA:EMULATE} may give an indicator. It will output an intermediate language of one the families of \textit{NV\_gpu\_programX} (where $X$ stands for the version). TEMP variables are registers. Since \textit{NV\_gpu\_programX} is a family of intermediate languages there is however no guarentees these are not further optimized. All other metrics discussed are readily available for OpenGL.

\textbf{Block size:} In general the block size should be chosen as small as possible to not hit resource limits. Though we in general had better experiences starting with a block size of 64 as can be seen in the Occupancy calculator. Increasing the block size is useful in combination with shared memory usage. Shared memory access is typically 100x faster than uncached global memory accesses. The more threads of a block reuse shared memory data the faster the memory throughput will be. Therefore, operators which involve filtering (also MPM-transfers) can use this behavior.

\textbf{Multiple elements per thread:} Merging threads and simultaneously reducing block size can reduce register pressure if the algorithm allows it. Otherwise merging sequential and parallel work improves performance heavily. Since the block size is limited by the GPU large amounts of data need to be split in tree structures to process in parallel. One can not avoid an indirection of writing-back and loading again from global memory. Therefore algorithms often improve much further from more sequential work to avoid this indirection.

\textbf{Loop unrolling:} Foremost loop unrolling reduces instructions and can thus speed up compute-bound work. Moreover loop unrolling might help the compiler reorder instructions and thus execute code faster. Instructions benefiting greatly from this are global loads. \cite{NVIDIA:PEAK_PERFORMANCE} shows at the example of motion blur and ray-marching that this greatly can increase throughput. Whereas the motion blur example uses a fixed amount of samples and therefore is easily unrollable, ray-marching is a dynamic process. Dynamic loops can however be partially unrolled. In the case of ray-marching this would for instance fetch two samples at once, although it might process only one. In short, this merges two consecutive loop iterations. The latency of the second instruction might be partly hidden by the first one and thus speed up the work load. Mind however that loop unrolling can increase register pressure. More registers need to store and process the relevant code, where a dynamic loop reuses the same registers for any iteration.

>>>>

There is no way to find out the exact number of registers an OpenGL shader uses. The program object assembly output of \textit{nvemulate} \cite{NVIDIA:EMULATE} may give an indicator. It will output an intermediate language of one the families of \textit{NV\_gpu\_programX} (where $X$ stands for the version). TEMP variables are registers. Since \textit{NV\_gpu\_programX} is a family of intermediate languages, there is however no guarentees these are not further optimized. Otherwise close monitoring of the occupancy is recommended. All other metrics discussed are readily available for OpenGL.

\textbf{Block size:} In general the block size should be chosen as small as possible to not hit resource limits. One should start however with a block size of at least 64 as can be seen in the Occupancy calculator. Increasing the block size is useful in combination with shared memory usage. Shared memory accesses are typically faster than global memory accesses. The more threads of a block reuse shared memory data the faster the memory throughput will be. Therefore, operators which involve filtering (also MPM-transfers) can use this behavior.

\textbf{Multiple elements per thread:} Merging threads and simultaneously reducing block size can reduce register pressure, if the algorithm allows it. Otherwise merging sequential and parallel work improves I/O latency hiding. Since the block size is limited by the GPU, some algorithms need to be split in tree structures to be processed. One can not avoid an indirection of writing back and loading again from global memory. Therefore algorithms often improve much further from more sequential work to avoid this indirection.

\textbf{Loop unrolling:} Foremost, loop unrolling reduces instructions and can thus speed up compute-bound work. Moreover, loop unrolling might help the compiler reorder instructions and thus execute code faster. Instructions benefiting greatly from this are global loads. \cite{NVIDIA:PEAK_PERFORMANCE} shows at the example of motion blur and ray-marching that this greatly can increase throughput. Whereas the motion blur example uses a fixed amount of samples and therefore is easily unrollable, ray-marching is a dynamic process. Dynamic loops can however be partially unrolled. In the case of ray-marching this would for instance fetch two samples at once although it might process only one. In short, this merges two consecutive loop iterations. The latency of the second instruction might be partly hidden by the first one and thus speed up the work load. Mind however that loop unrolling can increase register pressure. More registers need to stored and process the relevant code where a dynamic loop reuses the same registers for any iteration.
